---
title: 'lab 05: Predicting MHV Change'
author: "Ricky Duran"
date: "`r format(Sys.time(), '%B %d, %Y')`"
highlight: github
theme: cayman
Output:
  prettydoc::html_pretty: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, fig.width=10 )

library( dplyr )
library( knitr )
library( pander )
library( stargazer )
library( scales )
library( psych )

set.seed( 1234 )

# set stargazer type to text for 
# previewing in RMD docs but
# convert to type HTML when knitting
# (next code chunk)

s.type <- "text"  
```

```{r}
###################################
#
#     STARGAZER SETTINGS
#
###################################

# DO NOT RUN CHUNK UNLESS KNITTING:
# changes table formats to html
# before rendering RMD docs

#s.type <- "html"
```

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt)
    
    test <- cor.test(x,y)
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE,
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " "))
    
    text(0.5, 0.5, txt, cex = 1.5 )
    text(.7, .8, Signif, cex=cex, col=2)
}

panel.smooth <- function (x, y, col = par("col"), bg = NA, pch = par("pch"), 
    cex = 0.5, col.smooth = "red", span = 2/3, iter = 3, ...) 
{
    points(x, y, pch = 19, col = gray(0.7,0.2), bg = bg, cex = cex)
    ok <- is.finite(x) & is.finite(y)
    if (any(ok)) 
        lines(stats::lowess(x[ok], y[ok], f = span, iter = iter), 
            col = col.smooth, lwd=2, ...)
}

# custom plot
jplot <- function( x1, x2, lab1="", lab2="", draw.line=T, ... )
{

    plot( x1, x2,
          pch=19, 
          col=gray(0.6, alpha = 0.2), 
          cex=0.5,  
          bty = "n",
          xlab=lab1, 
          ylab=lab2, cex.lab=1.5,
        ... )

    if( draw.line==T ){ 
        ok <- is.finite(x1) & is.finite(x2)
        lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}
```

# Part 1: Data

Similar to your previous lab, create a dataset that includes 2000 and 2010 census variables drop all rural census tracts.

Create a variable that measures the growth of median home value from 2000 to 2010.

Omit cases that have a median home value less than $10,000 in 2000.

Omit cases with growth rates above 200%.

```{r}
d1 <- readRDS( "C:/Users/rduran3/Dropbox (ASU)/MPP/CPP 528 - Data Sciences III/lab05/data/LTDB-2000.rds" )
d2 <- readRDS( "C:/Users/rduran3/Dropbox (ASU)/MPP/CPP 528 - Data Sciences III/lab05/data/LTDB-2010.rds" )
md <- readRDS( "C:/Users/rduran3/Dropbox (ASU)/MPP/CPP 528 - Data Sciences III/lab05/data/LTDB-META-DATA.rds" )

d1 <- select( d1, - year )
d2 <- select( d2, - year )

d <- merge( d1, d2, by="tractid" )
d <- merge( d, md, by="tractid" )

d <- filter( d, urban == "urban" )

d.full <- d  # keep a copy so you don't have to reload 

d <- d.full  # reset to the original dataset

d <- select( d, tractid, 
             mhmval00, mhmval12, 
             hinc00, 
             hu00, vac00, own00, rent00, h30old00,
             empclf00, clf00, unemp00, prof00,  
             dpov00, npov00,
             ag25up00, hs00, col00, 
             pop00.x, nhwht00, nhblk00, hisp00, asian00,
             cbsa, cbsaname )

d <- 
  d %>%
  mutate( p.white = 100 * nhwht00 / pop00.x,
          p.black = 100 * nhblk00 / pop00.x,
          p.hisp = 100 * hisp00 / pop00.x, 
          p.asian = 100 * asian00 / pop00.x,
          p.hs = 100 * (hs00+col00) / ag25up00,
          p.col = 100 * col00 / ag25up00,
          p.prof = 100 * prof00 / empclf00,
          p.unemp = 100 * unemp00 / clf00,
          p.vacant = 100 * vac00 / hu00,
          mhv.change.00.to.10 = mhmval12 - mhmval00,
          p.mhv.change = 100 * (mhmval12 - mhmval00) / mhmval00,
          pov.rate = 100 * npov00 / dpov00 )


# adjust 2000 home values for inflation 
mhv.00 <- d$mhmval00 * 1.28855  
mhv.10 <- d$mhmval12

# change in MHV in dollars
mhv.change <- mhv.10 - mhv.00


# drop low 2000 median home values
# to avoid unrealistic growth rates.
#
# tracts with homes that cost less than
# $10,000 are outliers
mhv.00[ mhv.00 < 1000 ] <- NA

# change in MHV in percent
mhv.growth <- 100 * ( mhv.change / mhv.00 )

mhv.growth[ mhv.growth > 200 ] <- NA

d$mhv.00 <- mhv.00
d$mhv.10 <- mhv.10
d$mhv.change <- mhv.change
d$mhv.growth <- mhv.growth

pander(head(d %>% select(tractid, mhv.00, mhv.10, mhv.change, mhv.growth)))

(names(d))
```

# Part 2: Community Health

Reconstruct your three community health scales from Lab 02.

Run three bivariate regressions predicting MHV growth with each community health metric, and one regression with all three metrics together.

**The 3 instruments that I examined in Lab 2 were:**

**Professional Employees (p.prof):** As a measure of individual wellbing, this variable allows us to examine the economic strenth of an area. Since professsional level employment gives an individual a greater degree of financial security (higher payrates) and stability (typically include additional benefits, such as insurance), knowing the proportion of a community that possesses professional employment, allows us to understand the economic security of that community.
**Foreign Born Community Member (p.unemp):** As a measure of individual "wellbing" struggle, this variable allows us to examine the vulnerability of an area. Since unemployed individuals often face the barrier of financial strain, knowing the proportion of a community that is made up of unemployed community memebers, allows us to understand the vulnerability of that community.
**College Educated Individuals (p.col):** Having a college education allows us to determine, not only the degree of professional knowledge a person may possess, but can also be an indicator of a persons possession of soft skills and 'grit' or perciverance. A higher degree of community memebers with a college education may give us an idea of community health. 

## Determining the Reliability of our Indicies

By examining these indicies, we are able to determine the correlation and Cronbach's Alpha Score (Reliability) for this relationship. 

### Selected Variable Correlation:

Below we can see a visualization that allows us to compare correlations between variables. Noticing a high collelation between Professional Employment and College Educated Individuals, and a low correlation between thos variables and being forign-born. Although not all variables have strong corelations, I would like to point out that they all relationships are statistically significant. 

```{r}
df <- dplyr::select( d, 'p.prof', 'p.unemp', 'p.col', 'mhv.growth' )
pairs( df, lower.panel=panel.smooth, upper.panel=panel.cor )
```

### Selected Variable Alpha Score

```{r}
a1 <- psych::alpha( df, check.keys=TRUE )$total$raw_alpha
a1
```

With an Cronbach's Alpha Score of 71.82% we are able to say that this relationship is reliable.

### Descriptive Statistics of Metrics.

Below you will find the summary statistics for the variables being analized:

```{r}
summary (df) %>% pander()
```

```{r}
reg.data <- d

reg.data$mhv.growth[ reg.data$mhv.growth > 200 ] <- NA
reg.data$p.prof <- log10( reg.data$p.prof + 1 )
reg.data$p.unemp <- log10( reg.data$p.unemp + 1 )
reg.data$p.col <- log10( reg.data$p.col + 1 )

m1 <- lm( mhv.growth ~  p.prof, data=reg.data )
m2 <- lm( mhv.growth ~  p.unemp, data=reg.data )
m3 <- lm( mhv.growth ~  p.col, data=reg.data )
m4 <- lm( mhv.growth ~  p.prof + p.unemp +p.col, data=reg.data )


stargazer( m1, m2, m3, m4,
           type=s.type, digits=2,
           omit.stat = c("rsq","f") )
```

**Which metric, if any, best predicts change in home value? Does this fit with your expectations?**

**ANSWER:** All three metrics have a significant correlation with change in home value (p<0.01), although the metric that has the most distinct difference would be unemployment. 

# Part 3: Predict MHV Change

Now select at least three census variables that you feel will be good predictors of change in MHV between 2000 and 2010.

Run the model while including metro-level fixed effects (cbsa name or FIPS). Make sure you check for variable skew and multicollearity and adjust accordingly.

```{r}
d.reg <- d

reg.data <- d

d.reg$mhv.growth[ reg.data$mhv.growth > 200 ] <- NA
d.reg$p.prof <- log10( reg.data$p.prof + 1 )
d.reg$p.unemp <- log10( reg.data$p.unemp + 1 )
d.reg$p.col <- log10( reg.data$p.col + 1 )

# average growth in median home value for the city
d.reg <- 
  d.reg %>%
  group_by( cbsaname ) %>%
  mutate( metro.mhv.growth = 100 * median( mhv.growth, na.rm=T ) ) %>%
  ungroup() 

m1 <- lm( mhv.growth ~ p.prof, data=reg.data )
m2 <- lm( mhv.growth ~ p.unemp, data=reg.data )
m3 <- lm( mhv.growth ~ p.col, data=reg.data )
m4 <- lm( mhv.growth ~ p.prof + p.unemp + p.col + metro.mhv.growth, data=d.reg )
m5 <- lm( mhv.growth ~ p.prof + p.unemp + p.col + cbsa, data=d.reg )

stargazer( m1, m2, m3, m4, m5,
           type=s.type, digits=2,
           omit.stat = c("rsq","f"),
           omit="cbsa",
           add.lines = list(c("Metro Fixed Effects:", "NO", "NO", "No", "No","YES")) )
```

**What are the results? Which factor was most important? Did it meet your expectations? Were there any variables that were not significant that you expected to be? Explain your findings to a general audience.**

**Answer:** When running the regression with a cbsa fixed-effects, we notice that the correlation between change in home value and precent professional employees becomes less significant while home values correlgation with precent college graduates actually becomes mroe segnificant. All role of each role is still significant (p<0.05), although the variable that is highly significnt and with a highest rate of change would be unemployement, once again.


-------
<br>

Analysis Created By: Ricky Duran
For: CPP 528 - Data Sciences for Public Service III
Created On: "`r format(Sys.time(), '%B %d, %Y')`"

<br>
