---
title: "Lab 06 - Adding Federal Program Data"
author: "Jason Eastman"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( echo=TRUE, message=FALSE, warning=FALSE, fig.width=10 )
```

# Libraries

```{r}
library( dplyr )
library( knitr )
library( pander )
library( stargazer )
library( scales )
library( ggplot2 )
library( gridExtra )

set.seed( 1234 )

# For Stargazer
s.type <- "text"
```

# Add Federal Program Data to Model

## Load and Merge Datasets

```{r}
# Load datasets
d1 <- readRDS( "data/rodeo/LTDB-2000.rds" )
d2 <- readRDS( "data/rodeo/LTDB-2010.rds" )
md <- readRDS( "data/rodeo/LTDB-META-DATA.rds" )

d1 <- select( d1, - year )
d2 <- select( d2, - year )

# Merge datasets
d <- merge( d1, d2, by="tractid" )
census_data <- merge( d, md, by="tractid" )

```

## Add Federal Data

```{r}
URL1 <- "https://raw.githubusercontent.com/DS4PS/cpp-528-spr-2020/master/labs/data/raw/NMTC/nmtc-sheet-01.csv"
nmtc <- read.csv( URL1, stringsAsFactors=F )

URL <- "https://raw.githubusercontent.com/DS4PS/cpp-528-spr-2020/master/labs/data/raw/LIHTC/LIHTCPUB.csv"
lihtc <- read.csv( URL, stringsAsFactors=F )
```

## Clean Up Census Data

```{r}
census_data$id2 <- gsub( "[^0-9]", "", census_data$tractid )

census_data$id2 <- as.numeric( census_data$id2 )
```

## Aggregate Fed data

### LIHTC

```{r}
lihtc.dollars <-
  lihtc %>% 
  filter( yr_alloc >= 2000 & yr_alloc <= 2010 ) %>%
  group_by( fips2010 ) %>%
  summarize( num.lihtc = n(), lihtc.total = sum( allocamt, na.rm=T ) )

head( lihtc.dollars )
```

### NMTC

```{r}
# need to convert from currency to numeric
# current format: 
# head( nmtc$QLICI.Amount )
# [1] "$300,000.00 "   "$1,008,750.00 " "$977,000.00 "

# remove dollar sign and commas
nmtc$amount <- gsub( "[,$]", "", nmtc$QLICI.Amount )

# head(  nmtc$amount  )
# "300000.00 "  "1008750.00 " "977000.00 "

# convert characters to numeric 
nmtc$amount <- as.numeric( nmtc$amount ) %>% round(0)

# head(  nmtc$amount  )
# [1]  300000 1008750  977000

nmtc.dollars <- 
  nmtc %>% 
  filter( Origination.Year >= 2000 & Origination.Year <= 2010 ) %>%
  group_by( X2010.Census.Tract ) %>% 
  summarize( num.nmtc=n(), nmtc.total = sum( amount, na.rm=T ) )

head( nmtc.dollars )
```

## Merge Census and Tax Policy

```{r}
census_data <- merge( census_data, nmtc.dollars, by.x="id2", by.y="X2010.Census.Tract", all.x=T )
census_data <- merge( census_data, lihtc.dollars, by.x="id2", by.y="fips2010", all.x=T )

# recode tracts that had no grants from NA to 0

census_data$num.nmtc[ is.na(census_data$num.nmtc) ] <- 0
census_data$nmtc.total[ is.na(census_data$nmtc.total) ] <- 0

census_data$num.lihtc[ is.na(census_data$num.lihtc) ] <- 0 
census_data$lihtc.total[ is.na(census_data$lihtc.total) ] <- 0
```

## Remove Rural Districts

```{r}
census_data <- filter( census_data, urban == "urban" )
```

## Save Data into a Master

```{r}
data_master <- census_data
```

## Create New MHV Variables

```{r}
census_data <- data_master

 # adjust 2000 home values for inflation 
mhv.00 <- census_data$mhmval00 * 1.28855  
mhv.10 <- census_data$mhmval12

# change in MHV in dollars
mhv.change <- mhv.10 - mhv.00

# drop low 2000 median home values
# to avoid unrealistic growth rates.
#
# tracts with homes that cost less than
# $10,000 are outliers
mhv.00[ mhv.00 < 1000 ] <- NA

# change in MHV in percent
mhv.growth <- 100 * ( mhv.change / mhv.00 )

# drop high growth rate above 200%
# to avoid unrealistic growth rates.
#
# tracts with growth rates that are
# above 200% are outliers
mhv.growth[ mhv.growth > 200 ] <- NA

# Add variables into dataset
census_data$mhv.00 <- mhv.00
census_data$mhv.10 <- mhv.10
census_data$mhv.change <- mhv.change
census_data$mhv.growth <- mhv.growth 

# Peek at new variables
peek_at_data <- census_data %>%
  select(tractid, mhv.00, mhv.10, mhv.change, mhv.growth)

head(peek_at_data)
```

## Calculate New Metrics

```{r}
# Create subset , keeping only values we want
census_data <- select( census_data, 
             
             tractid, cbsa, cbsaname,            # ids / units of analysis
             
             mhv.00, mhv.10, mhv.change, mhv.growth,    # home value 
             
             hinc00, hu00, own00, rent00,        # ses
             hinc12, hu10, own10, rent10,
             
             empclf00, clf00, unemp00, prof00, flabf00,   # employment 
             empclf12, clf12, unemp12, prof12, flabf12,
             
             dpov00, npov00,                     # poverty
             dpov12, npov12,
             
             ag5up00, ag25up00, hs00, col00, lep00,             # education 
             ag5up12, ag25up12, hs12, col12, lep12,
             
             pop00.x, nhwht00, nhblk00, hisp00, asian00,   # race
             pop10, nhwht10, nhblk10, hisp10, asian10,
             
             num.nmtc, nmtc.total,              # tax policy data
             num.lihtc, lihtc.total             # aggregated by census tract
             ) # end select

census_data <- 
  census_data %>%
  mutate( 
          # 2000 variables
          p.white.00 = 100 * nhwht00 / pop00.x,
          p.black.00 = 100 * nhblk00 / pop00.x,
          p.hisp.00 = 100 * hisp00 / pop00.x, 
          p.asian.00 = 100 * asian00 / pop00.x,
          p.col.edu.00 = 100 * col00 / ag25up00,
          p.prof.00 = 100 * prof00 / empclf00,
          p.unemp.00 = 100 * unemp00 / clf00,
          pov.rate.00 = 100 * npov00 / dpov00,
          
          # 2010 variables
          p.white.10 = 100 * nhwht10 / pop10,
          p.black.10 = 100 * nhblk10 / pop10,
          p.hisp.10 = 100 * hisp10 / pop10, 
          p.asian.10 = 100 * asian10 / pop10,
          p.col.edu.10 = 100 * col12 / ag25up12,
          p.prof.10 = 100 * prof12 / empclf12,
          p.unemp.10 = 100 * unemp12 / clf12,
          pov.rate.10 = 100 * npov12 / dpov12,

          # More Chosen Variables added for 2000
          p.lep.00 = 100 * ( lep00 / ( ag5up00 + 0.01 )),
          p.flabf.00 = 100 * ( flabf00 / (clf00 + 0.01 )),
          
          # More Chosen Variables added for 2010
          p.lep.10 = 100 * ( lep12 / ( ag5up12 + 0.01 )),
          p.flabf.10 = 100 * ( flabf12 / (clf12 + 0.01 ))
  )

census_data$p.unemp.00 <- log( census_data$p.unemp.00 + 1)
census_data$p.lep.00 <- log( census_data$p.lep.00 + 1)
census_data$p.unemp.10 <- log( census_data$p.unemp.10 + 1)
census_data$p.lep.10 <- log( census_data$p.lep.10 + 1)


# inflation adjust income 
census_data$hinc00 <- 1.28855 * census_data$hinc00

# Create new variables
census_data <- 
  census_data %>%
  group_by( cbsaname ) %>%
  mutate( # metro rank of home value in 2000
          metro.mhv.pct.00 = ntile( mhv.00, 100 ),
          # metro rank of home value in 2010
          metro.mhv.pct.10 = ntile( mhv.10, 100 ),
          # median pay for metro area 2000
          metro.median.pay.00 = median( hinc00, na.rm=T ),
          # median pay for metro area 2010
          metro.median.pay.10 = median( hinc12, na.rm=T ),
          # tract rank in metro area for diversity (% non-white)
          metro.diversity.rank.00 = ntile( (100-p.white.00), 100 ),
          # metro total population 2000
          metro.pop.total.00 = sum( pop00.x, na.rm=T ),
          # metro total population 2010
          metro.pop.total.10 = sum( pop10, na.rm=T ),
          
          # More added variables
          
          # tract rank in metro area for unemployment in 2000
          metro.unemp.rank.00 = ntile( p.unemp.00, 100 ),
          # tract rank in metro area for unemployment in 2010
          metro.unemp.rank.10 = ntile( p.unemp.10, 100 ),
          # tract rank in metro area for Low English Speaking in 2000
          metro.lep.rank.00 = ntile( p.lep.00, 100 ),
          # tract rank in metro area for Low English Speaking in 2010
          metro.lep.rank.10 = ntile( p.lep.10, 100 ),
          # tract rank in metro area for Female Labor Force in 2000
          metro.flabf.rank.00 = ntile( p.flabf.00, 100 ),
          # tract rank in metro area for Female Labor Force in 2010
          metro.flabf.rank.10 = ntile( p.flabf.10, 100 )
          ) %>%
  ungroup() %>%
    mutate( # change in relative rank of mhv in metro area 2000-2010
          metro.mhv.pctile.change = metro.mhv.pct.10 - metro.mhv.pct.00,
          # growth in ave pay in metro
          metro.pay.change = metro.median.pay.10 - metro.median.pay.00,
          # metro population growth 2000-2010
          metro.pop.growth =
            ( metro.pop.total.10 - metro.pop.total.00 ) / metro.pop.total.00,
          # increase in the proportion of whites in tract 
          increase.p.white = p.white.10 - p.white.00,
          
          # More Added Variables
          
          # Growth in Unemployment from 2000 to 2010
          unemp.growth = p.unemp.10 - p.unemp.00,
          # Growth in Low English Speaking people from 2000 to 2010
          lep.growth = p.lep.10 - p.lep.00,
          # Growth in Female Labor Force from 2000 to 2010
          flabf.growth = p.flabf.10 - p.flabf.00
          )

# Round the entire dataset to 2 digits
is.num <- sapply(census_data, is.numeric)
census_data[is.num] <- lapply(census_data[is.num], round, 2)

# Create a true/false code for recipient tracts
census_data$LIHTC <- ifelse( census_data$num.lihtc > 0, "YES", "NO" )
census_data$NMTC <- ifelse( census_data$num.nmtc > 0, "YES", "NO" )

head(census_data)
```

# Diff in Diff Models

## Diff in Diff Model for Unemployment

**ANSWER:  See Code and output below for more info. From looking at Unemployment change between treatment and baseline groups, there is very little secualr trend (0.29). On the other hand, the treatment group shows a 10 percentage points decay than the baseline group.**

```{r}
y1 <- census_data$metro.unemp.rank.00
y2 <- census_data$metro.unemp.rank.10
treat <- as.numeric( census_data$num.nmtc > 0 )

d1 <- data.frame( y=y1, treat=treat, post=0 )
d2 <- data.frame( y=y2, treat=treat, post=1 )

diff_in_diff <- rbind( d1, d2 )

reg_diff <- lm( y ~ treat + post + treat*post, data=diff_in_diff )

summary( reg_diff ) 
```

## Diff in Diff Model for Low English Speaking People

**ANSWER:  See Code and output below for more info. We see that the baseline group shows very little to no growth rate (0.07), and moreover, the treatment growth shows a neglegent lesser growth (-2.77) than the baseline.**

```{r}
y1 <- census_data$metro.lep.rank.00
y2 <- census_data$metro.lep.rank.10
treat <- as.numeric( census_data$num.nmtc > 0 )

d1 <- data.frame( y=y1, treat=treat, post=0 )
d2 <- data.frame( y=y2, treat=treat, post=1 )

diff_in_diff <- rbind( d1, d2 )

reg_diff <- lm( y ~ treat + post + treat*post, data=diff_in_diff )

summary( reg_diff ) 
```

## Diff in Diff Model for Female Labor Force

**ANSWER: See Code and output below for more info. Here, both cases show neglegent to no growth, nor a difference between the baseline and the treatment groups.**

```{r}
y1 <- census_data$metro.flabf.rank.00
y2 <- census_data$metro.flabf.rank.10
treat <- as.numeric( census_data$num.nmtc > 0 )

d1 <- data.frame( y=y1, treat=treat, post=0 )
d2 <- data.frame( y=y2, treat=treat, post=1 )

diff_in_diff <- rbind( d1, d2 )

reg_diff <- lm( y ~ treat + post + treat*post, data=diff_in_diff )

summary( reg_diff ) 
```
