---
title: 'Lab 06: Adding Federal Program Data'
author: "Ricky Duran"
date: "`r format(Sys.time(), '%B %d, %Y')`"
highlight: github
theme: cayman
Output:
  prettydoc::html_pretty: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, fig.width=10 )

library( dplyr )
library( knitr )
library( pander )
library( stargazer )
library( scales )
library( ggplot2 )
library( gridExtra )

set.seed( 1234 )

# set stargazer type to text for 
# previewing in RMD docs but
# convert to type HTML when knitting
# (next code chunk)

#s.type <- "text"  
```

```{r}
###################################
#
#     STARGAZER SETTINGS
#
###################################

# DO NOT RUN CHUNK UNLESS KNITTING:
# changes table formats to html
# before rendering RMD docs

s.type <- "html"
```

# Overview
This last step in your project will walk you through adding data from two federal programs designed to help low-income

### About the NMTC Program

The NMTC Program enables economically distressed communities to leverage private investment capital by providing investors with a federal tax credit. All NMTC investments must meet statutory qualifications for their investors to be able to claim the tax credit. The vast majority of NMTC investments are made within statutorily defined “Low-Income Communities.” Low-Income Communities are census tracts with a poverty rate of 20 percent or greater, or a median family income at or below 80 percent of the applicable area median family income. In addition to investments located in Low-Income Communities, investments can qualify for NMTCs by using other statutory provisions designed to target certain areas or populations, including provisions for Rural Counties, and Low-Income Targeted Populations.

Through the first 15 application rounds of the NMTC Program, the CDFI Fund has made 1,178 awards, allocating a total of $57.5 billion in tax credit authority to CDEs through a competitive application process.

[data download website](https://www.cdfifund.gov/news-events/news/Pages/news-detail.aspx?NewsID=364&Category=Press%20Releases)

### About the LIHTC Program

The Low-Income Housing Tax Credit (LIHTC) is the most important resource for creating affordable housing in the United States today. The LIHTC database, created by HUD and available to the public since 1997, contains information on 47,511 projects and 3.13 million housing units placed in service between 1987 and 2017.

[data download website](https://lihtc.huduser.gov/)

# Data

Load your wrangled datasets and prepare your variables for analysis:

```{r}
d1 <- readRDS( "C:/Users/rduran3/Dropbox (ASU)/MPP/CPP 528 - Data Sciences III/lab05/data/LTDB-2000.rds" )
d2 <- readRDS( "C:/Users/rduran3/Dropbox (ASU)/MPP/CPP 528 - Data Sciences III/lab05/data/LTDB-2010.rds" )
md <- readRDS( "C:/Users/rduran3/Dropbox (ASU)/MPP/CPP 528 - Data Sciences III/lab05/data/LTDB-META-DATA.rds" )

d1 <- select( d1, - year )
d2 <- select( d2, - year )

d <- merge( d1, d2, by="tractid" )
d <- merge( d, md, by="tractid" )
```

# Add Federal Data

```{r}
URL1 <- "https://raw.githubusercontent.com/DS4PS/cpp-528-spr-2020/master/labs/data/raw/NMTC/nmtc-sheet-01.csv"
nmtc <- read.csv( URL1, stringsAsFactors=F )
head( nmtc ) 
```

```{r}
URL <- "https://raw.githubusercontent.com/DS4PS/cpp-528-spr-2020/master/labs/data/raw/LIHTC/LIHTCPUB.csv"
lihtc <- read.csv( URL, stringsAsFactors=F )
head( lihtc )
```

```{r}
# remove anything not a number from the string
d$id2 <- gsub( "[^0-9]", "", d$tractid )
head( d$id2 )
```

```{r}
# check NMTC census tract IDs

length( unique( nmtc$X2010.Census.Tract ) )
```

```{r}
length( intersect( nmtc$X2010.Census.Tract, d$id2 ) )
```

```{r}
# fix IDs so they are match
d$id2 <- as.numeric( d$id2 )

length( intersect( nmtc$X2010.Census.Tract, d$id2 ) )
```

```{r}
# all but 13 of projects have a matched census tract 


# check LIHTC census tract IDs

length( unique( lihtc$fips2010 ) )
```

```{r}
length( intersect( lihtc$fips2010, d$id2 ) )
```

```{r}
# all but 62 projects out of 21,000 have a matched census tract 
```

# Aggregate Fed Data

We can’t merger the dataset directly because there are multiple projects per census tract. We need to aggregate the project data first:

### LIHTC:

```{r}
lihtc.dollars <-
  lihtc %>% 
  filter( yr_alloc >= 2000 & yr_alloc <= 2010 ) %>%
  group_by( fips2010 ) %>%
  summarize( num.lihtc = n(), lihtc.total = sum( allocamt, na.rm=T ) )

head( lihtc.dollars )
```

### NMTC:

```{r}
# need to convert from currency to numeric
# current format: 
# head( nmtc$QLICI.Amount )
# [1] "$300,000.00 "   "$1,008,750.00 " "$977,000.00 "

# remove dollar sign and commas
nmtc$amount <- gsub( "[,$]", "", nmtc$QLICI.Amount )

# head(  nmtc$amount  )
# "300000.00 "  "1008750.00 " "977000.00 "

# convert characters to numeric 
nmtc$amount <- as.numeric( nmtc$amount ) %>% round(0)

# head(  nmtc$amount  )
# [1]  300000 1008750  977000

nmtc.dollars <- 
  nmtc %>% 
  filter( Origination.Year >= 2000 & Origination.Year <= 2010 ) %>%
  group_by( X2010.Census.Tract ) %>% 
  summarize( num.nmtc=n(), nmtc.total = sum( amount, na.rm=T ) )

head( nmtc.dollars )
```

# Merch Census and Tax Policy Data

```{r}
d <- merge( d, nmtc.dollars, by.x="id2", by.y="X2010.Census.Tract", all.x=T )
d <- merge( d, lihtc.dollars, by.x="id2", by.y="fips2010", all.x=T )

# recode tracts that had no grants from NA to 0

d$num.nmtc[ is.na(d$num.nmtc) ] <- 0
d$nmtc.total[ is.na(d$nmtc.total) ] <- 0

d$num.lihtc[ is.na(d$num.lihtc) ] <- 0 
d$lihtc.total[ is.na(d$lihtc.total) ] <- 0
```

# Remove Rural Districs

```{r}
table( d$urban )
```

```{r}
d <- filter( d, urban == "urban" )
```

How many tracts receive one or both programs:

```{r}
# total number of tracts: 59,722
length( unique( d$tractid ) )
```

```{r}
sum( d$num.lihtc > 0 )
```


```{r}
sum( d$num.nmtc > 0 )
```


```{r}
sum( d$num.lihtc > 0 & d$nmtc > 0 ) 
```

# Create New Variables

```{r}
d.full <- d  # keep a copy so you don't have to reload 
```

You will want to create more variables, but these examples will get you started.

```{r}
d <- d.full  # reset to the original dataset

# adjust 2000 home values for inflation 
mhv.00 <- d$mhmval00 * 1.28855  
mhv.10 <- d$mhmval12

# change in MHV in dollars
mhv.change <- mhv.10 - mhv.00


# drop low 2000 median home values
# to avoid unrealistic growth rates.
#
# tracts with homes that cost less than
# $10,000 are outliers
# approximately 200 out of 59,000 cases 
sum( mhv.00 < 10000 ) 
```

```{r}
mhv.00[ mhv.00 < 10000 ] <- NA

# change in MHV in percent
mhv.growth <- 100 * ( mhv.change / mhv.00 )
summary( mhv.growth )
```

```{r}
d$mhv.00 <- mhv.00
d$mhv.10 <- mhv.10
d$mhv.change <- mhv.change
d$mhv.growth <- mhv.growth 



d <- select( d, 
             
             tractid, cbsa, cbsaname,            # ids / units of analysis
             
             mhv.00, mhv.10, mhv.change, mhv.growth,    # home value 
             
             hinc00, hu00, own00, rent00,        # ses
             hinc12, hu10, own10, rent10,
             
             empclf00, clf00, unemp00, prof00,   # employment 
             empclf12, clf12, unemp12, prof12,
             
             dpov00, npov00,                     # poverty
             dpov12, npov12,
             
             ag25up00, hs00, col00,              # education 
             ag25up12, hs12, col12,
             
             pop00.x, nhwht00, nhblk00, hisp00, asian00,   # race
             pop10, nhwht10, nhblk10, hisp10, asian10,
             
             num.nmtc, nmtc.total,              # tax policy data
             num.lihtc, lihtc.total             # aggregated by census tract
             
          ) # end select


d <- 
  d %>%
  mutate( 
          # 2000 variables
          p.white.00 = 100 * nhwht00 / pop00.x,
          p.black.00 = 100 * nhblk00 / pop00.x,
          p.hisp.00 = 100 * hisp00 / pop00.x, 
          p.asian.00 = 100 * asian00 / pop00.x,
          p.hs.edu.00 = 100 * (hs00+col00) / ag25up00,
          p.col.edu.00 = 100 * col00 / ag25up00,
          p.prof.00 = 100 * prof00 / empclf00,
          p.unemp.00 = 100 * unemp00 / clf00,
          pov.rate.00 = 100 * npov00 / dpov00,
          
          # 2010 variables
          p.white.10 = 100 * nhwht10 / pop10,
          p.black.10 = 100 * nhblk10 / pop10,
          p.hisp.10 = 100 * hisp10 / pop10, 
          p.asian.10 = 100 * asian10 / pop10,
          p.hs.edu.10 = 100 * (hs12+col12) / ag25up12,
          p.col.edu.10 = 100 * col12 / ag25up12,
          p.prof.10 = 100 * prof12 / empclf12,
          p.unemp.10 = 100 * unemp12 / clf12,
          pov.rate.10 = 100 * npov12 / dpov12 )

# inflation adjust income 
d$hinc00 <- 1.28855 * d$hinc00

d <-
  d %>%
  group_by( cbsaname ) %>%
  mutate( # metro rank of home value in 2000
          metro.mhv.pct.00 = ntile( mhv.00, 100 ),
          # metro rank of home value in 2010
          metro.mhv.pct.10 = ntile( mhv.10, 100 ),
          # median pay for metro area 2000
          metro.median.pay.00 = median( hinc00, na.rm=T ),
          # median pay for metro area 2010
          metro.median.pay.10 = median( hinc12, na.rm=T ),
          # tract rank in metro area for diversity (% non-white)
          metro.diversity.rank.00 = ntile( (100-p.white.00), 100 ),
          # metro total population 2000
          metro.pop.total.00 = sum( pop00.x, na.rm=T ),
          # metro total population 2010
          metro.pop.total.10 = sum( pop10, na.rm=T ) ) %>%
  ungroup() %>%
  mutate( # change in relative rank of mhv in metro area 2000-2010
          metro.mhv.pctile.change = metro.mhv.pct.10 - metro.mhv.pct.00,
          # growth in ave pay in metro
          metro.pay.change = metro.median.pay.10 - metro.median.pay.00,
          # metro population growth 2000-2010
          metro.pop.growth = ( metro.pop.total.10 - metro.pop.total.00 ) / metro.pop.total.00,
          # increase in the proportion of whites in tract 
          increase.p.white = p.white.10 - p.white.00,
          # increase in the proportion of professionals in tract 
          increase.p.prof = p.prof.10 - p.prof.00,
          # increase in the proportion of unemployed community memebers in tract 
          increase.p.unemp = p.unemp.10 - p.unemp.00,
          # increase in the proportion of college educated community memebers in tract 
          increase.p.col.edu = p.col.edu.10 - p.col.edu.00
          )

d.master <- d
names(d)
```

# Use this script to add federal program data to your models.

Use the difference-in-difference framework introduced here to estimate the impact of each program.

- Use your baseline model predicting tract change from Lab 05 as the starting point.
- Create a log-linear diff-in-diff model following these steps and add your control variables from Lab 05.

```{r}
# Create a true/false code for recipient tracts
d$LIHTC <- ifelse( d$num.lihtc > 0, "YES", "NO" )
d$NMTC <- ifelse( d$num.nmtc > 0, "YES", "NO" )
```

### Profesionals

```{r}
plot1 <-
  ggplot( d, aes(x=p.prof.00, fill=NMTC )) +
  geom_density(alpha=0.4) + 
  ggtitle("2000 Professional Composition Comparison of \nRecipient and Non-Recipient Communities")

plot2 <-
  ggplot( d, aes(x=p.prof.00, fill=LIHTC)) +
  geom_density(alpha=0.4) +
  ggtitle("2000 Professionals Composition Comparison of \nRecipient and Non-Recipient Communities")

grid.arrange( plot1, plot2, nrow = 1 )
```

### Unemployment

```{r}
plot3 <-
  ggplot( d, aes(x=p.unemp.00, fill=NMTC )) +
  geom_density(alpha=0.4) + 
  ggtitle("2000 Unemployment Rate Comparison of \nRecipient and Non-Recipient Communities")

plot4 <-
  ggplot( d, aes(x=p.unemp.00, fill=LIHTC)) +
  geom_density(alpha=0.4) +
  ggtitle("2000 Unemployement Rate Comparison of \nRecipient and Non-Recipient Communities")

grid.arrange( plot3, plot4, nrow = 1 )
```

### College Education

```{r}
plot5 <-
  ggplot( d, aes(x=p.col.edu.00, fill=NMTC )) +
  geom_density(alpha=0.4) + 
  ggtitle("2000 Proportion College Educated Comparison of \nRecipient and Non-Recipient Communities")

plot6 <-
  ggplot( d, aes(x=p.col.edu.00, fill=LIHTC)) +
  geom_density(alpha=0.4) +
  ggtitle("2000 Proportion College Educated Comparison of \nRecipient and Non-Recipient Communities")

grid.arrange( plot5, plot6, nrow = 1 )
```

### Home Values

```{r}
plot7 <- 
  ggplot( d, aes(x=log10(mhv.00), fill=NMTC )) +
  geom_density( alpha=0.4 ) +
  ggtitle("2000 Median Home Value Comparison of \nRecipient and Non-Recipient Communities")

plot8 <- 
  ggplot( d, aes(x=log10(mhv.00), fill=LIHTC )) +
  geom_density( alpha=0.4 )  +
  ggtitle("2000 Median Home Value Comparison of \nRecipient and Non-Recipient Communities")

grid.arrange( plot7, plot8, nrow = 1 )
```

# Use this script to add federal program data to your models.

Use the difference-in-difference framework introduced here to estimate the impact of each program.

- Use your baseline model predicting tract change from Lab 05 as the starting point.
- Create a log-linear diff-in-diff model following these steps and add your control variables from Lab 05.

## Evaluating Program Impact

Difference-in-Difference Model:

Back to the diff-in-diff option. It turns out there is a relatively easy fix to the challenge of estimating growth from two points in time. It is possible because of the magic of log functions.

In the regression context, logs change our interpretation of slopes from a one-unit change in X being associated with a B-unit change in Y, to a one-unit change in X being associated with a growth rate of B for Y.

So back to the home value problem. A is worth $200k and B $100k. They both grow at the same rate. Home A increases in value by $20k, and home B by $10k.

Once logged, however, note the important approximation:

log( A[t=2] ) - log( A[t=1] ) is approximately equal to ( (A2-A1)/A1 ) or the growth rate.

So if we log the home value in the diff-in-diff models then we can calculate growth rates as:

log(C2) - log(C1) = growth rate of comparison / secular market trend

log(T2) - log(T1) = growth rate of treatment

log(C1) = B0

log(C2) = B0 + B1

secular growth rate = log(C2) - log(C1) = (B0 + B1) - B0 = B1

B1 represents the defaul growht rate of home values for the comparison group.

The important coefficient, B3, would then represent the growth rate of the treatment group above the secular growth rate, or the growth rate necessary to generate a home value of T2 if we start from a value of T1.

```{r}
y1 <- log1p( d$p.prof.00 )
y2 <- log1p( d$p.prof.10 )
treat <- as.numeric( d$num.nmtc > 0 )

d1 <- data.frame( y=y1, treat=treat, post=0 )
d2 <- data.frame( y=y2, treat=treat, post=1 )

d3 <- rbind( d1, d2 )

m <- lm( y ~ treat + post + treat*post, data=d3 )

summary( m ) 
```

```{r}
y1 <- log1p( d$p.unemp.00 )
y2 <- log1p( d$p.unemp.10 )
treat <- as.numeric( d$num.nmtc > 0 )

d1 <- data.frame( y=y1, treat=treat, post=0 )
d2 <- data.frame( y=y2, treat=treat, post=1 )

d3 <- rbind( d1, d2 )

m <- lm( y ~ treat + post + treat*post, data=d3 )

summary( m ) 
```

```{r}
y1 <- log1p( d$p.col.edu.00 )
y2 <- log1p( d$p.col.edu.10 )
treat <- as.numeric( d$num.nmtc > 0 )

d1 <- data.frame( y=y1, treat=treat, post=0 )
d2 <- data.frame( y=y2, treat=treat, post=1 )

d3 <- rbind( d1, d2 )

m <- lm( y ~ treat + post + treat*post, data=d3 )

summary( m ) 
```

**Report and interpret the main results from the models. Are the programs effective at catalyzing neighborhood improvement? We are assuming median home value captures the increased market demand that reflects the desirability of improvements.**

**ANSWER:** 
Proportion Professionals: the the secular trend was -5.58% loss. For this program the treatment group grew 31.11 percentage points more than the baseline group.

Unemployment Rate: the the secular trend was 49.15% growth. For this program the treatment group decreased 30.78 percentage points faster than the baseline group.

Proportion College Educated: the the secular trend was 14.68% growth. For this program the treatment group increased 13.70 percentage points faster than the baseline group.

# REFLECTION:

**How can we test the parallel lines assumption in this model? We know that growth rates change significantly between periods. The market for urban homes from 1990-2000 looks very different from the market in 2000 to 2010.**

I would say that we would have to look at the population we are examining. Are we looking at the community? We would be able to tell if the # of long time residents decreased, or if the reacial make up chages... Esentially we would be looking to see whether we have impacted to community member originally in the area, and did not cause a case of gentrification.

-------
<br>

Analysis Created By: Ricky Duran
For: CPP 528 - Data Sciences for Public Service III
Created On: "`r format(Sys.time(), '%B %d, %Y')`"

<br>
