---
title: "Lab-04-Gentry"
author: "Katie Gentry"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: cerulean
    highlight: haddock
---

## Load packages and data needed to complete analysis of filtering and cleaning data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library( dplyr )
library( knitr )
library( pander )
library( stargazer )
library( scales )
```

```{r}
d1 <- readRDS( "~/R/win-library/analytic_project/cpp-528-spr-2020-group-01/data/rodeo/LTDB-1990.rds" )
d2 <- readRDS( "~/R/win-library/analytic_project/cpp-528-spr-2020-group-01/data/rodeo/LTDB-2000.rds" )

d3 <- readRDS( "~/R/win-library/analytic_project/cpp-528-spr-2020-group-01/data/rodeo/LTDB-2010.rds" )

md <- readRDS( "~/R/win-library/analytic_project/cpp-528-spr-2020-group-01/data/rodeo/LTDB-META-DATA.rds" )

# check to make sure we are not losing 
# or gaining observations in the merge
nrow( d1 ) 
```

```{r}
d1 <- select( d1, - year )
d2 <- select( d2, - year )
d3 <- select( d3, - year )

d <- merge( d1, d2, by="tractid" )
d <- merge( d, md, by="tractid" )

dat <- merge( d2, d3, by="tractid" )
dat <- merge( dat, md, by="tractid" )

```

## Filter to just urban districts. 

```{r}
table(d$urban.x)

d <- filter( d, urban.x == "urban" )
dat <- filter (dat, urban.x == "urban")
```

## Find variables in both files and identify shared data points.

```{r}
# find variables that are in both files
compare_dfs <- function( df1, df2 )
{
  # use regular expressions to remove numeric suffixes 
  var.names.1 <- names( df1 )
  var.names.1 <- gsub( "[.][xy]$", "", var.names.1 )
  var.names.1 <- gsub( "[0-9]{2}$", "", var.names.1 )
  
  var.names.2 <- names( df2 )
  var.names.2 <- gsub( "[.][xy]$", "", var.names.2 )
  var.names.2 <- gsub( "[0-9]{2}$", "", var.names.2 )
  
  shared <- intersect( var.names.1, var.names.2 ) %>% sort()
  print( "SHARED VARIABLES:")
  print( shared )
  
  not.shared <- c( setdiff( var.names.1, var.names.2 ),
                   setdiff( var.names.2, var.names.1 ) ) %>% sort()
  
  print( "NOT SHARED:" )
  print( not.shared )
  
  d.vars1 <- data.frame( type="shared", variables=shared, stringsAsFactors=F )
  d.vars2 <- data.frame( type="not shared", variables=not.shared, stringsAsFactors=F )
  dd <- rbind( d.vars1, d.vars2 )
  
  return( dd )
}

vars <- compare_dfs( df1=d1, df2=d2 )
```

```{r}
head( vars )
```

# Part 1

## Create data subset for analysis

```{r}
d.full <- d
d.data <- dat
```

```{r}
d <- d.full  # store original in case you need to reset anything

d <- select( d, tractid, mhmval90, mhmval00, hinc90, 
             hu90, own90, rent90,  
             empclf90, clf90, unemp90, prof90,  
             dpov90, npov90,
             ag25up90, hs90, col90, 
             pop90.x, nhwht90, nhblk90, hisp90, asian90,
             cbsa.x, cbsaname.x)

d <- 
  d %>%
  mutate( p.white = 100 * nhwht90 / pop90.x,
          p.black = 100 * nhblk90 / pop90.x,
          p.hisp = 100 * hisp90 / pop90.x, 
          p.asian = 100 * asian90 / pop90.x,
          p.hs = 100 * (hs90+col90) / ag25up90,
          p.col = 100 * col90 / ag25up90,
          p.prof = 100 * prof90 / empclf90,
          p.unemp = 100 * unemp90 / clf90,
          pov.rate = 100 * npov90 / dpov90 )
```

## View summary statistics in Stargazer

```{r}
stargazer( d, 
           type="text", 
           digits=0,
           summary.stat = c("min", "p25","median","mean","p75","max") )
```

## Median Home Value Exploration

```{r}
# adjust 2000 home values for inflation 
mhv.90 <- d$mhmval90 * 1.34522  
mhv.00 <- d$mhmval00

mhv.change <- mhv.00 - mhv.90

df <- data.frame( MedianHomeValue1990=mhv.90, 
                  MedianHomeValue2000=mhv.00, 
                  Change.90.to.00=mhv.change )

stargazer( df, 
           type="text", 
           digits=0, 
           summary.stat = c("min", "p25","median","mean","p75","max") )
```

```{r}
hist( mhv.change/1000, breaks=500, 
      xlim=c(-100,500), yaxt="n", xaxt="n",
      xlab="Thousand of US Dollars (adjusted to 2000)", cex.lab=1.5,
      ylab="", main="Change in Median Home Value 1990 to 2000",
      col="gray20", border="white" )

axis( side=1, at=seq( from=-100, to=500, by=100 ), 
      labels=paste0( "$", seq( from=-100, to=500, by=100 ), "k" ) )
        
mean.x <- mean( mhv.change/1000, na.rm=T )
abline( v=mean.x, col="darkorange", lwd=2, lty=2 )
text( x=200, y=1500, 
      labels=paste0( "Mean = ", dollar( round(1000*mean.x,0)) ), 
      col="darkorange", cex=1.8, pos=3 )

median.x <- median( mhv.change/1000, na.rm=T )
abline( v=median.x, col="dodgerblue", lwd=2, lty=2 )
text( x=200, y=2000, 
      labels=paste0( "Median = ", dollar( round(1000*median.x,0)) ), 
      col="dodgerblue", cex=1.8, pos=3 )
```

```{r}
# function to control plot() formatting 
jplot <- function( x1, x2, lab1="", lab2="", draw.line=T, ... )
{

    plot( x1, x2,
          pch=19, 
          col=gray(0.6, alpha = 0.2), 
          cex=2.5,  
          bty = "n",
          xlab=lab1, 
          ylab=lab2, cex.lab=1.5,
        ... )

    if( draw.line==T ){ 
        ok <- is.finite(x1) & is.finite(x2)
        lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}
```

```{r}
layout.matrix <- matrix( c( 1,3,
                            2,3 ), 
                nrow=2, ncol=2, byrow=T )

layout( mat = layout.matrix,
        heights = c(2,2), # Heights of the two rows
        widths =  c(3,4)) # Widths of the two columns

# layout.show(3)

par( mar=c(4,0,0,2) )

hist( mhv.90/1000, breaks=50, 
      xlim=c(-200,800), yaxt="n", xaxt="n",
      xlab="", cex.lab=1,
      ylab="", main="",
      col="darkslateblue", border="white" )

axis( side=1, at=seq( from=0, to=1000, by=100 ), 
      labels=paste0( "$", seq( from=0, to=1000, by=100 ), "k" ) )

abline( v=seq(0,1000,100), lty=2, col="gray80" )

text( 550, 4000, labels="Median Home \nValue in 1990", 
      col="darkslateblue", cex=1.8 )



hist( mhv.00/1000, breaks=50, 
      xlim=c(-200,800), yaxt="n", xaxt="n",
      xlab="", cex.lab=1,
      ylab="", main="",
      col="darkslateblue", border="white" )

abline( v=seq(0,1000, 100 ), lty=2, col="gray80" )

text( 550, 3500, labels="Median Home \nValue in 2000", 
      col="darkslateblue", cex=1.8 )

axis( side=1, at=seq( from=0, to=1000, by=100 ), 
      labels=paste0( "$", seq( from=0, to=1000, by=100 ), "k" ) )


# data reduction - filter 1,000 observations

df <- data.frame( v90=mhv.90/1000, v00=mhv.00/1000 )
df <- sample_n( df, 1000 )

par( mar=c(4,5,3,2) )

jplot( df$v90, df$v00, 
       lab1="MHV in 1990", lab2="MHV in 2000",
       xlim=c(0,1000), ylim=c(0,1000),
       axes=F )

abline( a=0, b=1, lty=2, col="gray" )
axis( side=1, at=seq( from=0, to=1000, by=200 ), 
      labels=paste0( "$", seq( from=0, to=1000, by=200 ), "k" ) )
axis( side=2, at=seq( from=0, to=1000, by=200 ), 
      labels=paste0( "$", seq( from=0, to=1000, by=200 ), "k" ) )
```

## Change in Median Home Value 1990-2000

```{r}
# small initial values are skewing percentages
#
# an average home value below $10k is really low -
# these must be mostly vacant lots?

# interpretation is hard if there were no homes in 2000
# and thus an artificially low MHV. i don't trust cases
# that go from homes worth $10k to regular value
# because it is more likely errors in data or noise
# than meaningful variance 
#
# quick filter to remove all of the problematic obs
# but need to go back and see which cases are problematic


mhv.90[ mhv.90 < 1000 ] <- NA
pct.change <- mhv.change / mhv.90
summary( pct.change )
```

```{r}
sum( pct.change > 5, na.rm=T )
```

```{r}
# preview tracts with large increases in home values 
# to see if increases make sense 

d %>% 
  filter( pct.change > 5 ) %>% 
  head() 
```

```{r}
hg <-
hist( pct.change, breaks=2000, 
      xlim=c(-1,2), yaxt="n", xaxt="n",
      xlab="", cex.main=1.5,
      ylab="", main="Growth in Home Value by Census Tract 1990 to 2000",
      col="gray40", border="white" )

axis( side=1, at=seq( from=-1, to=2, by=0.5 ), 
      labels=paste0( seq( from=-100, to=200, by=50 ), "%" ) )

ymax <- max( hg$count )
        
mean.x <- mean( pct.change, na.rm=T )
abline( v=mean.x, col="darkorange", lwd=2, lty=2 )
text( x=1, y=(0.5*ymax), 
      labels=paste0( "Mean = ", round(100*mean.x,0), "%"), 
      col="darkorange", cex=1.8, pos=4 )

median.x <- median( pct.change, na.rm=T )
abline( v=median.x, col="dodgerblue", lwd=2, lty=2 )
text( x=1, y=(0.6*ymax), 
      labels=paste0( "Median = ", round(100*median.x,0), "%"), 
      col="dodgerblue", cex=1.8, pos=4 )
```

## Growth Rates by Area

```{r}
d$mhv.change <- mhv.change 
d$pct.change <- pct.change
d$mhv.90 <- mhv.90
d$mhv.00 <- mhv.00

d %>%
  group_by( cbsaname.x ) %>%
  summarize( ave.change = median( mhv.change, na.rm=T ),
             ave.change.d = dollar( round(ave.change,0) ),
             growth = 100 * median( pct.change, na.rm=T ) ) %>%
  ungroup() %>%
  arrange( - growth ) %>%
  select( - ave.change ) %>% 
  head( 25 ) %>%
  pander()
```

# Part 1 Questions

**Between 1990 and 2000, Median Home Values either decline or rise very insignificantly compared to the inflation rate. Comparatively, home values between 2000 and 2010, rose significantly compared to the inflation rate. The authors suggest that lower income, non-educated, and typically minority individuals left these urban spaces due to the rising home prices and the lack of labor opportunities.**

# Part 2

```{r}
#unemployment rates
unemp.00 <- d.data$unemp00
unemp.10 <- d.data$unemp12

unemp.change <- unemp.10 - unemp.00

df <- data.frame( Unemployed2000= unemp.00, 
                  Unemployed2010= unemp.10,
                  Change.00to.10= unemp.change)

stargazer( df, 
           type="text", 
           digits=0, 
           summary.stat = c("min", "p25","median","mean","p75","max") )
```

```{r}
hist( unemp.change, breaks=500, 
      xlim=c(-100,500), yaxt="n", xaxt="n",
      xlab="Number Unemployed", cex.lab=1.5,
      ylab="", main="Change in Unemployment 2000 to 2010",
      col="gray20", border="white" )

axis( side=1, at=seq( from=-120, to=500, by=10), 
      labels=paste0(seq( from=-120, to=500, by=10 )) )
        
mean.x <- mean( unemp.change, na.rm=T )
abline( v=mean.x, col="darkorange", lwd=2, lty=2 )
text( x=200, y=1500, 
      labels=paste0( "Mean = ", ( round(mean.x,0)) ), 
      col="darkorange", cex=1.8, pos=3 )

median.x <- median( unemp.change, na.rm=T )
abline( v=median.x, col="dodgerblue", lwd=2, lty=2 )
text( x=200, y=2000, 
      labels=paste0( "Median = ", ( round(median.x,0)) ), 
      col="dodgerblue", cex=1.8, pos=3 )
```

```{r}
# function to control plot() formatting 
jplot <- function( x1, x2, lab1="", lab2="", draw.line=T, ... )
{

    plot( x1, x2,
          pch=19, 
          col=gray(0.6, alpha = 0.2), 
          cex=2.5,  
          bty = "n",
          xlab=lab1, 
          ylab=lab2, cex.lab=1.5,
        ... )

    if( draw.line==T ){ 
        ok <- is.finite(x1) & is.finite(x2)
        lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}
```

```{r}
layout.matrix <- matrix( c( 1,3,
                            2,3 ), 
                nrow=2, ncol=2, byrow=T )

layout( mat = layout.matrix,
        heights = c(2,2), # Heights of the two rows
        widths =  c(3,4)) # Widths of the two columns

# layout.show(3)

par( mar=c(4,0,0,2) )

hist( unemp.00, breaks=50, 
      xlim=c(-120,500), yaxt="n", xaxt="n",
      xlab="", cex.lab=1,
      ylab="", main="",
      col="darkslateblue", border="white" )

axis( side=1, at=seq( from=-100, to=500, by=10 ), 
      labels=paste0( seq( from=-100, to=500, by=10 )) )

abline( v=seq(-100,500,100), lty=2, col="gray80" )

text( 550, 4000, labels="Unemployment in 2000", 
      col="darkslateblue", cex=1.8 )

hist( unemp.10, breaks=50, 
      xlim=c(-120,500), yaxt="n", xaxt="n",
      xlab="", cex.lab=1,
      ylab="", main="",
      col="darkslateblue", border="white" )

axis( side=1, at=seq( from=-100, to=500, by=10 ), 
      labels=paste0( seq( from=-100, to=500, by=10 )) )

abline( v=seq(-100,500,100), lty=2, col="gray80" )

text( 550, 4000, labels="Unemployment in 2010", 
      col="darkslateblue", cex=1.8 )

# data reduction - filter 1,000 observations

df <- data.frame( v00=unemp.00, v10=unemp.10)
df <- sample_n( df, 1000 )

par( mar=c(4,5,3,2) )

jplot( df$v00, df$v10, 
       lab1="Unemp in 2000", lab2="Unemp in 2010",
       xlim=c(0,500), ylim=c(0,500),
       axes=F )

abline( a=0, b=1, lty=2, col="gray" )
axis( side=1, at=seq( from=0, to=500, by=50 ), 
      labels=paste0( seq( from=0, to=500, by=50 )) )
axis( side=1, at=seq( from=0, to=500, by=50 ), 
      labels=paste0( seq( from=0, to=500, by=50 )) )
```

```{r}
unemp.pct.change <- unemp.change / (unemp.00+.01)
summary( unemp.pct.change )
```

```{r}
unempg <-
hist( unemp.pct.change, breaks=1000, 
      xlim=c(-1,2), yaxt="n", xaxt="n",
      xlab="", cex.main=1.5,
      ylab="", main="Growth in Unemployment by Census Tract 2000 to 2010",
      col="gray40", border="white" )

axis( side=1, at=seq( from=-1, to=2, by=0.5 ), 
      labels=paste0( seq( from=-100, to=200, by=50 ), "%" ) )

ymax <- max( hg$count )
        
mean.x <- mean( unemp.pct.change, na.rm=T )
abline( v=mean.x, col="darkorange", lwd=2, lty=2 )
text( x=1, y=(0.5*ymax), 
      labels=paste0( "Mean = ", round(mean.x,0), "%"), 
      col="darkorange", cex=1.8, pos=4 )

median.x <- median( unemp.pct.change, na.rm=T )
abline( v=median.x, col="dodgerblue", lwd=2, lty=2 )
text( x=1, y=(1*ymax), 
      labels=paste0( "Median = ", round(median.x,0), "%"), 
      col="dodgerblue", cex=1.8, pos=4 )
```

**We see that unemployment increased which is unexpectant given that home values also increased. Due to the 2007-2009 recession, we know that the unemployment rate was still greater in 2010 than in 2000. By census tract, unemployment grew 42% which in this case is a negative.**

**On the contrary, we see an average increase in rent.**

```{r}
#renter rates
rent.00 <- d.data$rent00
rent.10 <- d.data$rent12

rent.change <- rent.10 - rent.00

df <- data.frame( Renters2000= rent.00, 
                  Renters2010= rent.10,
                  Change.00to.10= rent.change)

stargazer( df, 
           type="text", 
           digits=0, 
           summary.stat = c("min", "p25","median","mean","p75","max") )
```

```{r}
hist( rent.change, breaks=500, 
      xlim=c(-200,500), yaxt="n", xaxt="n",
      xlab="Rent", cex.lab=1.5,
      ylab="", main="Change in Rent 2000 to 2010",
      col="gray20", border="white" )

axis( side=1, at=seq( from=-200, to=500, by=50), 
      labels=paste0(seq( from=-200, to=500, by=50 )) )
        
mean.x <- mean( rent.change, na.rm=T )
abline( v=mean.x, col="darkorange", lwd=2, lty=2 )
text( x=200, y=1500, 
      labels=paste0( "Mean = ", ( round(mean.x,0)) ), 
      col="darkorange", cex=1.8, pos=3 )

median.x <- median( rent.change, na.rm=T )
abline( v=median.x, col="dodgerblue", lwd=2, lty=2 )
text( x=200, y=2000, 
      labels=paste0( "Median = ", ( round(median.x,0)) ), 
      col="dodgerblue", cex=1.8, pos=3 )
```

```{r}
# function to control plot() formatting 
jplot <- function( x1, x2, lab1="", lab2="", draw.line=T, ... )
{

    plot( x1, x2,
          pch=19, 
          col=gray(0.6, alpha = 0.2), 
          cex=2.5,  
          bty = "n",
          xlab=lab1, 
          ylab=lab2, cex.lab=1.5,
        ... )

    if( draw.line==T ){ 
        ok <- is.finite(x1) & is.finite(x2)
        lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}
```

```{r}
layout.matrix <- matrix( c( 1,3,
                            2,3 ), 
                nrow=2, ncol=2, byrow=T )

layout( mat = layout.matrix,
        heights = c(2,2), # Heights of the two rows
        widths =  c(3,4)) # Widths of the two columns

# layout.show(3)

par( mar=c(4,0,0,2) )

hist( rent.00, breaks=500, 
      xlim=c(0,800), yaxt="n", xaxt="n",
      xlab="", cex.lab=1,
      ylab="", main="",
      col="darkslateblue", border="white" )

axis( side=1, at=seq( from=0, to=1200, by=50 ), 
      labels=paste0( seq( from=0, to=1200, by=50 )) )

abline( v=seq(0,800,100), lty=2, col="gray80" )

text( 650, 1500, labels="Rent in 2000", 
      col="darkslateblue", cex=1.4 )

hist( rent.10, breaks=500, 
      xlim=c(0,800), yaxt="n", xaxt="n",
      xlab="", cex.lab=1,
      ylab="", main="",
      col="darkslateblue", border="white" )

axis( side=1, at=seq( from=0, to=1200, by=50 ), 
      labels=paste0( seq( from=0, to=1200, by=50 )) )

abline( v=seq(0,800,100), lty=2, col="gray80" )

text( 650, 1500, labels="Rent in 2010", 
      col="darkslateblue", cex=1.4 )

# data reduction - filter 1,000 observations

df <- data.frame( v00=rent.00, v10=rent.10)
df <- sample_n( df, 1000 )

par( mar=c(4,5,3,2) )

jplot( df$v00, df$v10, 
       lab1="Rent in 2000", lab2="Rent in 2010",
       xlim=c(-100,800), ylim=c(-100,800),
       axes=F )

abline( a=0, b=1, lty=2, col="gray" )
axis( side=1, at=seq( from=-100, to=800, by=50 ), 
      labels=paste0( seq( from=-100, to=800, by=50 )) )
axis( side=1, at=seq( from=-100, to=800, by=50 ), 
      labels=paste0( seq( from=-100, to=800, by=50 )) )
```

```{r}
rent.pct.change <- rent.change / (rent.00+.01)
summary( rent.pct.change )
```

```{r}
rentg <-
hist( rent.pct.change, breaks=1000, 
      xlim=c(-1,2), yaxt="n", xaxt="n",
      xlab="", cex.main=1.5,
      ylab="", main="Growth in Rent by Census Tract 2000 to 2010",
      col="gray40", border="white" )

axis( side=1, at=seq( from=-1, to=2, by=0.5 ), 
      labels=paste0( seq( from=-100, to=200, by=50 ), "%" ) )

ymax <- max( rentg$count )
        
mean.x <- mean( rent.pct.change, na.rm=T )
abline( v=mean.x, col="darkorange", lwd=2, lty=2 )
text( x=1, y=(0.5*ymax), 
      labels=paste0( "Mean = ", round(mean.x,0), "%"), 
      col="darkorange", cex=1.8, pos=4 )

median.x <- median( rent.pct.change, na.rm=T )
abline( v=median.x, col="dodgerblue", lwd=2, lty=2 )
text( x=1, y=(1*ymax), 
      labels=paste0( "Median = ", round(median.x,0), "%"), 
      col="dodgerblue", cex=1.8, pos=4 )
```


# Part 3

```{r}
d4 <- dat
```

```{r}
# adjust 2000 home values for inflation 
mhv.00 <- d4$mhmval00 * 1.28855  
mhv.10 <- d4$mhmval12

mhv.change <- mhv.10 - mhv.00

# small initial values are skewing percentages
#
# an average home value below $10k is really low -
# these must be mostly vacant lots?

mhv.00[ mhv.00 < 1000 ] <- NA
pct.change <- 100 * ( mhv.change / mhv.00 )
summary( pct.change )
```

```{r}
d4$mhv.00 <- mhv.00
d4$mhv.10 <- mhv.10
d4$mhv.change <- mhv.change
d4$pct.change <- pct.change
```

```{r}
head( vars )
```

```{r}
d4 <- select( d4, 
             
             tractid, cbsa.x, cbsaname.x,            # ids / units of analysis
             
             mhv.00, mhv.10, mhv.change, pct.change,    # home value
             
             hinc00, hu00, own00, rent00,        # ses
             hinc12, hu10, own10, rent10,
             
             empclf00, clf00, unemp00, prof00,   # employment 
             empclf12, clf12, unemp12, prof12,
             
             dpov00, npov00,                     # poverty
             dpov12, npov12,
             
             ag25up00, hs00, col00,              # education 
             ag25up12, hs12, col12,
             
             pop00.x, nhwht00, nhblk00, hisp00, asian00,   # race
             pop10, nhwht10, nhblk10, hisp10, asian10
             
          ) # end select


d4 <- 
  d4 %>%
  mutate( 
          # 2000 variables
          p.white.00 = 100 * nhwht00 / pop00.x,
          p.black.00 = 100 * nhblk00 / pop00.x,
          p.hisp.00 = 100 * hisp00 / pop00.x, 
          p.asian.00 = 100 * asian00 / pop00.x,
          p.hs.edu.00 = 100 * (hs00+col00) / ag25up00,
          p.col.edu.00 = 100 * col00 / ag25up00,
          p.prof.00 = 100 * prof00 / empclf00,
          p.unemp.00 = 100 * unemp00 / clf00,
          pov.rate.00 = 100 * npov00 / dpov00,
          p.rent.00 = 100 * rent00 / hu00,
          
          # 2010 variables
          p.white.10 = 100 * nhwht10 / pop10,
          p.black.10 = 100 * nhblk10 / pop10,
          p.hisp.10 = 100 * hisp10 / pop10, 
          p.asian.10 = 100 * asian10 / pop10,
          p.hs.edu.10 = 100 * (hs12+col12) / ag25up12,
          p.col.edu.10 = 100 * col12 / ag25up12,
          p.prof.10 = 100 * prof12 / empclf12,
          p.unemp.10 = 100 * unemp12 / clf12,
          pov.rate.10 = 100 * npov12 / dpov12,
          p.rent.10 = 100 * rent10 / hu10 )
```

```{r}
d4 <-
  d4 %>%
  group_by( cbsaname.x ) %>%
  mutate( metro.mhv.pct.00 = ntile( mhv.00, 100 ),
          metro.mhv.pct.10 = ntile( mhv.10, 100 ),
          metro.median.pay.00 = median( hinc00, na.rm=T ),
          metro.median.pay.10 = median( hinc12, na.rm=T ),
          metro.race.rank.00 = ntile( (100-p.white.00), 100 ) ) %>%
  ungroup() %>%
  mutate( metro.mhv.pct.change = metro.mhv.pct.10 - metro.mhv.pct.00,
          pay.change = metro.median.pay.10 - metro.median.pay.00,
          race.change = p.white.10 - p.white.00,
          mhv.change = mhv.10 - mhv.00,
          pov.change = pov.rate.10 - pov.rate.00,
          p.unemp.change = p.unemp.10 - p.unemp.00,
          p.col.edu.change = p.col.edu.10 - p.col.edu.00,
          p.rent.change = p.rent.10 - p.rent.00)
```

```{r}
d5 <-           
  d4 %>%
  select( c( "tractid", "cbsa.x", "cbsaname.x",
             "mhv.00", "mhv.10", "mhv.change","pct.change",
          "p.white.00", "p.black.00", "p.hisp.00", "p.asian.00", 
          "p.hs.edu.00", "p.col.edu.00", "p.prof.00",  "p.unemp.00", 
          "pov.rate.00", "p.white.10", "p.black.10", "p.hisp.10", 
          "p.asian.10", "p.hs.edu.10", "p.col.edu.10", "p.prof.10", 
          "p.unemp.10", "pov.rate.10", "metro.mhv.pct.00", 
          "metro.mhv.pct.10", "metro.median.pay.00", "metro.median.pay.10", 
          "metro.mhv.pct.change", "pay.change", "race.change",
          "metro.race.rank.00", "pov.change", "p.unemp.change", "p.col.edu.change", "p.rent.change") ) 
```

```{r}
d5 <- data.frame(d5)
stargazer( d5, 
           type="text", 
           digits=0, 
           summary.stat = c("min", "p25","median","mean","p75","max") )
```

```{r}
# education status
# unemployment
# poverty
# median income
# % renters

# proportion of college educated increased by 3 percent
less.uneducated <- d5$p.col.edu.change > 3 

# proportion of unemployment decreased by 3 percent
less.unemployed <- d5$p.unemp.change > -3 

# proportion of poverty decreased by 3 percent
less.poverty <- d5$pov.change > -3 

# proportion of renters decreased
less.renters <- d5$p.rent.change > 0

# median income increased
more.money <- d5$pay.change > 0

# poverty in higher than normal rate
poor.2000 <- d5$pov.rate.00 > 9 

# median income for metros less than normal rate
income.2000 <- d5$metro.median.pay.00 < 43139

g.flag <- poor.2000 & income.2000 & less.uneducated & less.poverty & less.unemployed & less.renters & more.money

num.candidates <-  sum( poor.2000 & income.2000, na.rm=T )
num.gentrified <- sum( g.flag, na.rm=T )

num.gentrified 
```

```{r}
num.candidates
```

```{r}
num.gentrified / num.candidates
```

**If you define gentrification in this way, 17% of the tracts gentrified. I chose to define gentrifiation as more educated individuals present, less unemployment, less poverty, higher median income, and less renters. Theoretically, as renters become buyers due to higher income and less poverty is present, more gentrification is also happening.**

# Part 4

```{r}
library( geojsonio )  # read geoJSON map files from GitHub
library( sp )         # spatial data class sp for shapefiles
library( cartogram )  # spatial maps w/ tract size bias reduction
library( tmap )       # thematic maps
library( maptools )   # spatial object manipulation 
library( sf )         # 'simple features' flavor of shapefiles


# we have phoenix already packaged and on GitHub for easy load: 

github.url <- "https://raw.githubusercontent.com/DS4PS/cpp-529-master/master/data/phx_dorling.geojson"
phx <- geojson_read( x=github.url,  what="sp" )
plot( phx )
```

```{r}
# create small dataframe for the merge
df <- data.frame(  tractid=dat$tractid, 
        mhv.00,  mhv.10,  mhv.change,  pct.change  )

# create GEOID that matches GIS format

# create a geoID for merging by tract 
df$GEOID <- substr( df$tractid, 6, 18 )  # extract codes
df$GEOID <- gsub( "-", "", df$GEOID )    # remove hyphens
class( df$GEOID )
```

```{r}
head( df$GEOID )
```

```{r}
head( phx@data )  # sp class from GIS file, so data frame is located @data
```

```{r}
# merge census data with dorling map

nrow( phx ) # check dimensions
```

```{r}
phx <- merge( phx, df, by.x="GEOID", by.y="GEOID" )

# make sure they have not changed or 
# you are missing rows in your data frame
# or merging with the wrong ID
nrow( phx ) 
```

```{r}
phx <- spTransform( phx, CRS("+init=epsg:3395") )

bb <- st_bbox( c( xmin = -12519146, xmax = -12421368, 
                 ymax = 3965924, ymin = 3899074 ), 
               crs = st_crs("+init=epsg:3395")) 


tm_shape( phx, bbox=bb ) + 
  tm_polygons( col="mhv.00", n=10, style="quantile", palette="Spectral" ) +
  tm_layout( "Dorling Cartogram", title.position=c("right","top") )
```

```{r}
tm_shape( phx, bbox=bb ) + 
  tm_polygons( col="mhv.change", n=10, style="quantile", palette="Spectral" ) +
  tm_layout( "Dorling Cartogram", title.position=c("right","top") )
```

```{r}
tm_shape( phx, bbox=bb ) + 
  tm_polygons( col="pct.change", n=10, style="quantile", palette="Spectral" ) +
  tm_layout( "Dorling Cartogram", title.position=c("right","top") )
```
## Home Values

**The distribution of homes with higher values are located in Paradise Valley, Scottsdale, Gilbert, Chandler, and Ahwatukee. The lower priced value homes are located in West Phoenix, Glendale, and Mesa which is to be expected. The largest gains occur in Central Phoenix including Arcadia, Tempe, and North Phoenix. These homes were average in 2000 but had significant gains.**

```{r}
# create small dataframe for the merge
df1 <- data.frame(  tractid=d$tractid, 
        d5$p.col.edu.00, d5$p.col.edu.10, d5$p.col.edu.change)

# create GEOID that matches GIS format

# create a geoID for merging by tract 
df1$GEOID <- substr( df1$tractid, 6, 18 )  # extract codes
df1$GEOID <- gsub( "-", "", df1$GEOID )    # remove hyphens
class( df1$GEOID )
```

```{r}
head( df1$GEOID )
```

```{r}
head( phx@data )  # sp class from GIS file, so data frame is located @data
```

```{r}
# merge census data with dorling map

nrow( phx ) # check dimensions
```

```{r}
phx1 <- merge( phx, df1, by.x="GEOID", by.y="GEOID" )

# make sure they have not changed or 
# you are missing rows in your data frame
# or merging with the wrong ID
nrow( phx1 ) 
```

```{r}
phx1 <- spTransform( phx1, CRS("+init=epsg:3395") )

bb <- st_bbox( c( xmin = -12519146, xmax = -12421368, 
                 ymax = 3965924, ymin = 3899074 ), 
               crs = st_crs("+init=epsg:3395")) 


tm_shape( phx1, bbox=bb ) + 
  tm_polygons( col="college edu", n=2, style="quantile", palette="Spectral" ) +
  tm_layout( "Dorling Cartogram", title.position=c("right","top") )
```














