---
title: "Lab-06-Gentry"
author: "Katie Gentry"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: cerulean
    highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( echo=TRUE, message=FALSE, warning=FALSE, fig.width=10 )
```

# Libraries

```{r}
library( dplyr )
library( knitr )
library( pander )
library( stargazer )
library( scales )
library( ggplot2 )
library( gridExtra )
library( here )

set.seed( 1234 )
```

# Add Federal Program Data to Model

## Load and Merge Datasets

```{r}
# Load datasets
data.2000 <- readRDS( here("data/rodeo/LTDB-2000.rds" ))
data.2010 <- readRDS( here("data/rodeo/LTDB-2010.rds" ))
md <- readRDS( here("data/rodeo/LTDB-META-DATA.rds" ))

data.2000 <- select( data.2000, - year )
data.2010 <- select( data.2010, - year )

# Merge datasets
d <- merge( data.2000, data.2010, by="tractid" )
censusdata <- merge( d, md, by="tractid" )
```

## Add Federal Data

```{r}
URL1 <- "https://raw.githubusercontent.com/DS4PS/cpp-528-spr-2020/master/labs/data/raw/NMTC/nmtc-sheet-01.csv"
nmtc <- read.csv( URL1, stringsAsFactors=F )
```

```{r}
URL <- "https://raw.githubusercontent.com/DS4PS/cpp-528-spr-2020/master/labs/data/raw/LIHTC/LIHTCPUB.csv"
lihtc <- read.csv( URL, stringsAsFactors=F )
```

## Clean Up Census Data

```{r}
# remove anything not a number from the string
censusdata$id2 <- gsub( "[^0-9]", "", censusdata$tractid )
head( censusdata$id2 )
```

```{r}
# check NMTC census tract IDs

length( unique( nmtc$X2010.Census.Tract ) )
```

```{r}
length( intersect( nmtc$X2010.Census.Tract, censusdata$id2 ) )
```

```{r}
# fix IDs so they are match
censusdata$id2 <- as.numeric( censusdata$id2 )

length( intersect( nmtc$X2010.Census.Tract, censusdata$id2 ) )
```

```{r}
# all but 13 of projects have a matched census tract 

# check LIHTC census tract IDs

length( unique( lihtc$fips2010 ) )
```

```{r}
length( intersect( lihtc$fips2010, censusdata$id2 ) )
```

## We must aggregate data in order to merge data sets

```{r}
lihtc.dollars <-
  lihtc %>% 
  filter( yr_alloc >= 2000 & yr_alloc <= 2010 ) %>%
  group_by( fips2010 ) %>%
  summarize( num.lihtc = n(), lihtc.total = sum( allocamt, na.rm=T ) )

head( lihtc.dollars )
```

```{r}
# need to convert from currency to numeric
# current format: 
# head( nmtc$QLICI.Amount )
# [1] "$300,000.00 "   "$1,008,750.00 " "$977,000.00 "

# remove dollar sign and commas
nmtc$amount <- gsub( "[,$]", "", nmtc$QLICI.Amount )

# head(  nmtc$amount  )
# "300000.00 "  "1008750.00 " "977000.00 "

# convert characters to numeric 
nmtc$amount <- as.numeric( nmtc$amount ) %>% round(0)

# head(  nmtc$amount  )
# [1]  300000 1008750  977000

nmtc.dollars <- 
  nmtc %>% 
  filter( Origination.Year >= 2000 & Origination.Year <= 2010 ) %>%
  group_by( X2010.Census.Tract ) %>% 
  summarize( num.nmtc=n(), nmtc.total = sum( amount, na.rm=T ) )

head( nmtc.dollars )
```

## Merge Census data and Tax Policy data

```{r}
d <- merge( censusdata, nmtc.dollars, by.x="id2", by.y="X2010.Census.Tract", all.x=T )
d <- merge( d, lihtc.dollars, by.x="id2", by.y="fips2010", all.x=T )

# recode tracts that had no grants from NA to 0

d$num.nmtc[ is.na(d$num.nmtc) ] <- 0
d$nmtc.total[ is.na(d$nmtc.total) ] <- 0

d$num.lihtc[ is.na(d$num.lihtc) ] <- 0 
d$lihtc.total[ is.na(d$lihtc.total) ] <- 0
```

## Remove rural districts

```{r}
table( d$urban.x )
```

```{r}
# filter to just urban districts
d <- filter( d, urban.x == "urban" )
```

```{r}
#determine lihtc assistance tracts
sum( d$num.lihtc > 0 )
```

```{r}
#determine nmtc assistance tracts
sum( d$num.nmtc > 0 )
```

```{r}
#determine lihtc and nmtc tracts
sum( d$num.lihtc > 0 & d$nmtc > 0 ) 
```

## Create New Variables

```{r}
# pro tip when getting started
# and exploring a lot of variables
# while making changes to your dataset:

d.full <- d  
```

```{r}
d <- d.full 

# adjust 2000 home values for inflation 
mhv.00 <- d$mhmval00 * 1.28855  
mhv.10 <- d$mhmval12

# change in MHV in dollars
mhv.change <- mhv.10 - mhv.00


# drop low 2000 median home values
# to avoid unrealistic growth rates.
#
# tracts with homes that cost less than
# $10,000 are outliers
# approximately 200 out of 59,000 cases 
sum( mhv.00 < 10000 ) 
mhv.00[ mhv.00 < 10000 ] <- NA

# change in MHV in percent
mhv.growth <- 100 * ( mhv.change / mhv.00 )
summary( mhv.growth )
```
```{r}
head(d)
```


```{r}
d$mhv.00 <- mhv.00
d$mhv.10 <- mhv.10
d$mhv.change <- mhv.change
d$mhv.growth <- mhv.growth 



d <- select( d, 
             
             tractid, cbsa.x, cbsaname.x,            # ids / units of analysis
             
             mhv.00, mhv.10, mhv.change, mhv.growth,    # home value 
             
             hinc00, hu00, own00, rent00,        # ses
             hinc12, hu10, own10, rent10, 
             
             empclf00, clf00, unemp00, prof00, semp00,  # employment 
             empclf12, clf12, unemp12, prof12, semp12,
             
             dpov00, npov00,                     # poverty
             dpov12, npov12,
             
             ag25up00, hs00, col00,              # education 
             ag25up12, hs12, col12,
             
             pop00.x, nhwht00, nhblk00, hisp00, asian00,   # race
             pop10, nhwht10, nhblk10, hisp10, asian10,
             
             num.nmtc, nmtc.total,              # tax policy data
             num.lihtc, lihtc.total             # aggregated by census tract
             
          ) # end select

d <- 
  d %>%
  mutate( 
          # 2000 variables
          p.white.00 = 100 * nhwht00 / pop00.x,
          p.black.00 = 100 * nhblk00 / pop00.x,
          p.hisp.00 = 100 * hisp00 / pop00.x, 
          p.asian.00 = 100 * asian00 / pop00.x,
          p.hs.edu.00 = 100 * (hs00+col00) / ag25up00,
          p.col.edu.00 = 100 * col00 / ag25up00,
          p.prof.00 = 100 * prof00 / empclf00,
          p.unemp.00 = 100 * unemp00 / clf00,
          pov.rate.00 = 100 * npov00 / dpov00,
          p.selfemp.00 = 100 * semp00 / empclf00,
          
          # 2010 variables
          p.white.10 = 100 * nhwht10 / pop10,
          p.black.10 = 100 * nhblk10 / pop10,
          p.hisp.10 = 100 * hisp10 / pop10, 
          p.asian.10 = 100 * asian10 / pop10,
          p.hs.edu.10 = 100 * (hs12+col12) / ag25up12,
          p.col.edu.10 = 100 * col12 / ag25up12,
          p.prof.10 = 100 * prof12 / empclf12,
          p.unemp.10 = 100 * unemp12 / clf12,
          pov.rate.10 = 100 * npov12 / dpov12,
          p.selfemp.10 = 100 * semp12 / empclf12)

# inflation adjust income 
d$hinc00 <- 1.28855 * d$hinc00

d <-
  d %>%
  group_by( cbsaname.x ) %>%
  mutate( # metro rank of home value in 2000
          metro.mhv.pct.00 = ntile( mhv.00, 100 ),
          # metro rank of home value in 2010
          metro.mhv.pct.10 = ntile( mhv.10, 100 ),
          # median pay for metro area 2000
          metro.median.pay.00 = median( hinc00, na.rm=T ),
          # median pay for metro area 2010
          metro.median.pay.10 = median( hinc12, na.rm=T ),
          # tract rank in metro area for diversity (% non-white)
          metro.diversity.rank.00 = ntile( (100-p.white.00), 100 ),
          # metro total population 2000
          metro.pop.total.00 = sum( pop00.x, na.rm=T ),
          # metro total population 2010
          metro.pop.total.10 = sum( pop10, na.rm=T ) ) %>%
  ungroup() %>%
  mutate( # change in relative rank of mhv in metro area 2000-2010
          metro.mhv.pctile.change = metro.mhv.pct.10 - metro.mhv.pct.00,
          # growth in ave pay in metro
          metro.pay.change = metro.median.pay.10 - metro.median.pay.00,
          # metro population growth 2000-2010
          metro.pop.growth = ( metro.pop.total.10 - metro.pop.total.00 ) / metro.pop.total.00,
          # increase in the proportion of whites in tract 
          increase.p.white = p.white.10 - p.white.00,
          # increase in the proportion of self-employed in tract 
          increase.p.selfemp = p.selfemp.10 - p.selfemp.00,
          # increase in the proportion of unemployed community members in tract 
          increase.p.unemp = p.unemp.10 - p.unemp.00)

```

## Program Participation


```{r}
# Create a true/false code to determine recipient tracts
d$LIHTC <- ifelse( d$num.lihtc > 0, "YES", "NO" )
d$NMTC <- ifelse( d$num.nmtc > 0, "YES", "NO" )
```

## Compare Plots of Recipient and Non-Recipient Communities

### Self-Employment

```{r}
plot1 <-
  ggplot( d, aes(x=p.selfemp.00, fill=NMTC )) +
  geom_density(alpha=0.4) + 
  ggtitle("2000 Self-Employed \nComparison of Recipient and \nNon-Recipient Communities")

plot2 <-
  ggplot( d, aes(x=p.selfemp.00, fill=LIHTC)) +
  geom_density(alpha=0.4) +
  ggtitle("2000 Self-Employed \nComparison of Recipient and \nNon-Recipient Communities")

grid.arrange( plot1, plot2, nrow = 1 )
```

### Unemployment

```{r}
plot3 <-
  ggplot( d, aes(x=p.unemp.00, fill=NMTC )) +
  geom_density(alpha=0.4) + 
  ggtitle("2000 Unemployment Rate \nComparison of Recipient and \nNon-Recipient Communities")

plot4 <-
  ggplot( d, aes(x=p.unemp.00, fill=LIHTC)) +
  geom_density(alpha=0.4) +
  ggtitle("2000 Unemployement Rate \nComparison of Recipient and \nNon-Recipient Communities")

grid.arrange( plot3, plot4, nrow = 1 )
```

### Metro Median Pay

```{r}
plot5 <-
  ggplot( d, aes(x=metro.median.pay.00, fill=NMTC )) +
  geom_density(alpha=0.4) + 
  ggtitle("2000 Metro Median Pay \nComparison of Recipient and \nNon-Recipient Communities")

plot6 <-
  ggplot( d, aes(x=metro.median.pay.00, fill=LIHTC)) +
  geom_density(alpha=0.4) +
  ggtitle("2000 Metro Median Pay \nComparison of Recipient and \nNon-Recipient Communities")

grid.arrange( plot5, plot6, nrow = 1 )
```

### Home Values

```{r}
plot7 <- 
  ggplot( d, aes(x=log10(mhv.00), fill=NMTC )) +
  geom_density( alpha=0.4 ) +
  ggtitle("2000 Median Home Value \nComparison of Recipient and \nNon-Recipient Communities")

plot8 <- 
  ggplot( d, aes(x=log10(mhv.00), fill=LIHTC )) +
  geom_density( alpha=0.4 )  +
  ggtitle("2000 Median Home Value \nComparison of Recipient and \nNon-Recipient Communities")

grid.arrange( plot7, plot8, nrow = 1 )
```

## Evaluating Program Impact by Difference-in-Difference Models:

```{r}
y1 <- log1p( d$p.selfemp.00 )
y2 <- log1p( d$p.selfemp.10 )
treat <- as.numeric( d$num.nmtc > 0 )

d1 <- data.frame( y=y1, treat=treat, post=0 )
d2 <- data.frame( y=y2, treat=treat, post=1 )

d3 <- rbind( d1, d2 )

m <- lm( y ~ treat + post + treat*post, data=d3 )

summary( m ) 
```

**For Self-employed individuals, the secular trend was 4.17% loss. The treatment group grew 1.38% more than the baseline group so the programs were minimally effective at catalyzing neighborhood improvement.**

```{r}
y1 <- log1p( d$p.unemp.00 )
y2 <- log1p( d$p.unemp.10 )
treat <- as.numeric( d$num.nmtc > 0 )

d1 <- data.frame( y=y1, treat=treat, post=0 )
d2 <- data.frame( y=y2, treat=treat, post=1 )

d3 <- rbind( d1, d2 )

m <- lm( y ~ treat + post + treat*post, data=d3 )

summary( m ) 
```

**For unemployment rate, the secular trend was 49.1% growth. The treatment group was -30.8% lower than the baseline group so the programs were effective at catalyzing neighborhood improvement as there was lower unemployment.**

```{r}
y1 <- log1p( d$metro.median.pay.00)
y2 <- log1p( d$metro.median.pay.10)
treat <- as.numeric( d$num.nmtc > 0 )

d1 <- data.frame( y=y1, treat=treat, post=0 )
d2 <- data.frame( y=y2, treat=treat, post=1 )

d3 <- rbind( d1, d2 )

m <- lm( y ~ treat + post + treat*post, data=d3 )

summary( m ) 
```

**For Metro Median Pay, the secular trend was -4.24% loss. The treatment group grew 1.20% more than the baseline group so the programs were minimally effective at catalyzing neighborhood improvement.**
